
# **shiny-plotly**
Repo for training and experimentation in data interactivity and visualisation

## Notes

* Function in ```ui.R``` tells Shiny where to display objects. 
* Build the objects in ```server.R```. The code should go in the unnamed function ```function(input, output){}``` inside the ```shinyServer()``` function. 


<br>

### Lesson 4 - Reactive outputs

<br>

#### Reactive output as a two step process.

1. Add an **R object** to your user-interface with ```ui.R```.
2. Tell Shiny how to build the object in ```server.R```. The object will be reactive if the *code that builds it calls a* ***widget value***.

<br>

#### ```ui.R``` functions: turn R objects into output for your user-interface.

| Output function    | creates  |
|--------------------|----------|
| ```htmlOutput```         | raw HTML |
| ```imageOutput```        | image    |
| ```plotOutput```         | plot     |
| ```tableOutput```        | table    |
| ```textOutput```         | text     |
| ```uiOutput```           | raw HTML |
| ```verbatimTextOutput``` | text     |

*** Place the output function inside ```sidebarPanel``` or ```mainPanel``` in the ```ui.R``` script. ***

<br>

#### ```server.R``` functions: provide content to the UI
##### Create reactivity by connecting the ***values*** of ```input``` to the ***objects*** in ```output```. 

* Code should go in the unnamed function ```function(input, output){}``` inside the ```shinyServer()``` function. 
  + builds a list-like object named ```output``` that contains **all of the code needed to update the R objects in your app**. 
  + Each R object needs to **have its own entry** in the list.
    + Create an entry **matching the name of reactive element** created in ```ui.R``` by defining new element for ```output```.
    + Each entry to ```output``` should contain the output of one of Shiny’s ```render*``` functions, **corrresponding to the type** of reactive object being built.
  
  e.g. 
  
```{r eval = FALSE}
  # server.R

shinyServer(function(input, output) {

     output$text1 <- renderText({ 
          "You have selected this"
     })

  }
)
```


| render function | creates                                         |
|-----------------|-------------------------------------------------|
| ```renderImage```     | images (saved as a link to a source file)       |
| ```renderPlot```      | plots                                           |
| ```renderPrint```     | any printed output                              |
| ```renderTable```     | data frame, matrix, other table like structures |
| ```renderText```      | character strings                               |
| ```renderUI```        | a Shiny tag object or HTML                      |



Each ```render*``` function takes a single argument: an R expression surrounded by braces, ```{}```

<br>


#### Use widget values to build reactive content

* Make objects reactive by using ```input``` argument in ```render.R``` functions to call widget output values from UI.
      * name of ```render.R``` ```input$``` values must match ```inputId``` argument of desired widget output
e.g.

  
```{r eval = FALSE}
# server.R

shinyServer(
  function(input, output) {
  
    output$text1 <- renderText({ 
      paste("You have selected", input$var)
    })
    
  }
)
```
  
  


